---
title: 'Beyond the P-Value: Visualizing and Embracing Uncertainty'
output:
  html_document:
    df_print: paged
---

```{r setup, warning = F, message = F}
packages = c("infer",
             "dplyr",
             "nullabor",
             "ggplot2")

## Figure out which packages are not installed and install them
installed = vapply(packages, function(x) 
  length(find.package(x, quiet = TRUE)) > 0, 
  TRUE)

if(any(!installed)) {
  install.packages(packages[!installed])
}

## Load packages
done = lapply(packages, require, character.only = TRUE)


```


## P-value: Living in the world of the null hypothesis

I tried to bend a quarter to make it "unfair" and yield tails more often. Did I succeed?

$$ H_0: p = 0.5$$
$$H_A: p < 0.5$$

Let's say I flip it 5 times and got heads only once.  How unusual is that under the null hypothesis? We can easily simulate this in R using the `infer` package

```{r test_heads}
## Number of heads. Change this to see what happens!
number_heads = 4
n = 5
observed_p = number_heads/n

## Enter data
my_data = data.frame(flip = c(rep("T", n - number_heads), 
                               rep("H", number_heads)))

null_flips = my_data %>% 
  specify(response = flip, success = "H") %>% 
  hypothesize("point", p = 0.5) %>% 
  generate(reps = 1000, type = "simulate")

```

This simulates flipping a fair coin 5 times, and repeating the process 1000 times.  Here's the first three groups of 5:

```{r}
head(null_flips, 15)

```

Then we calculate the number of heads per flip and plot the distribution of sets of flips, and see where one heads actually falls on that distribution.

```{r}
null_proportions = null_flips %>% 
  calculate(stat = "prop")

visualise(null_proportions) + 
  shade_pvalue(observed_p, direction = "greater")
```

The p-value is just the proportions of times we got one or fewer heads:

```{r}
null_proportions %>% 
  get_pvalue(observed_p, direction = "greater")
```


There's a few great web apps for exploring this too.

For the coin flip situation, I recommend the 

## Experiment 1 data

`exp1.csv` contains 100 subjects who took the ESP test (tried to guess which curtain the picture was behind).  During each session, some of the pictures were erotic and some were non-erotic.

Variables are as follows:

  - *Session* Order in which participant was seen.
  - *Session_Type* (Added EB) Some participants saw different combinations of stimuli.  This is my best guess from Bem's writing about which was which.
  - *Num_Erotic* (Added EB) Number of erotic trials, varies by session type; my best guess.
  - *Num_Control* (Added EB) Number of control trials, varies by session type; my best guess.
  - *Erotic.Hits.PC* Percentage correct on erotic trials
  - *Control.Hits.PC* Percentage correct on control trials
  - *Stimulus.Seeking* 1-5 scale on how "stimulus-seeking" the participant is
  - *Date*
  - *StartTime*
  - *Session.Length*
  - *Participant.Sex*
  - *Participant.Age*
  - *ExpSex* ?

Below is code for reading in all trials, calculating the overall hit rate, and testing whether it is significantly different from chance:

$$ H_0 = \mu_\text{hit rate} = 50$$
$$ H_A = \mu_\text{hit rate} > 50$$
What did we actually see on average?
```{r}
# Original source: https://replicationindex.files.wordpress.com/2018/01/exp1.xlsx
# Some variables added as noted above.

exp1 = read.csv("exp1.csv", stringsAsFactors = FALSE)

exp1_all_hits = exp1 %>%
  mutate(All.Hits.PC = (Erotic.Hits.PC * Num_Erotic + Control.Hits.PC * Num_Control)/(Num_Erotic + Num_Control))

all_hits_mean = exp1_all_hits %>% 
    specify(response = All.Hits.PC) %>% 
    calculate(stat = "mean")

all_hits_mean
```

```{r}
null_all_hits = exp1_all_hits %>% 
  specify(response = All.Hits.PC) %>% 
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")

visualize(null_all_hits) + 
  shade_pvalue(all_hits_mean, direction = "greater")

```
```{r}
null_all_hits %>% 
  get_pvalue(all_hits_mean, direction = "greater")
```

We can also do a good old-fashioned t-test, for a similar result:

```{r}
t.test(exp1_all_hits$All.Hits.PC, mu = 50, alternative = "greater")
```

Can you find any significant findings here?

```{r}
t.test(exp1_all_hits$Erotic.Hits.PC, mu = 50, alternative = "greater")
```

```{r}
t.test(exp1_all_hits$Control.Hits.PC, mu = 50, alternative = "greater")
```



## Visual hypothesis testing -- the line-up

The dataset `mtcars`, included in R, contains the fuel consumption (`mpg`) and weight (`wt`) of 32 cars in 1974.  See `?mtcars` for more information.

Here, our question is: is there any visible relationship between `mpg` and `wt`?

We can do a *visual hypothesis test*. We mix up `mpg` to break the relationship between `mpg` and `wt` using `lineup` from the *`nullabor`* package.  (The example below is taken from the *`nullabor`* vignette.)

```{r}
d = lineup(null_permute("mpg"), mtcars)

ggplot(data=d, aes(x=mpg, y=wt)) + 
  geom_point() + 
  facet_wrap(~ .sample)

```

Can you tell which one is the real data?

We can reveal the position by accessing the `"pos"` attribute, using `attr(d, "pos")`.  There are also features to hide the position from a user by using `encrypt()`!

If one person correctly guesses the correct value, then the visual p-value to reject the null hypothesis is 0.05 -- they had a one in twenty chance of getting it correct.  If you ask more than one person and they all get it correct, that increases your ability to get a low visual p-value.  If 100 people are given the same 20 plots but only 15 get it correct, the p-value is still quite small:

```{r}
pvisual(15, 100, m = 20)
```

## Blinded analysis
We can take this idea a step further and do a *whole analysis* where somehow the true value has been blinded.

See `exp2_blinded.csv` for an example.

## Bayesian analysis

[Example app for Bayes](https://tellmi.psy.lmu.de/felix/BayesLessons/BayesianLesson1.Rmd)

## Resources

We're hear to help with any of this! [Research Methodology Consulting Center](http://cehd.umn.edu/research/consulting/)

And our friends—

[LATIS](https://cla.umn.edu/latis)

[IRSA](https://irsa.stat.umn.edu)

[Libraries](https://www.lib.umn.edu)

[DASH](https://dash.umn.edu)


