---
title: "Understanding Statistical Power and Avoiding P-Hacking"
output:
  html_document:
    df_print: paged
---


```{r setup, warning = F, message = F}
packages = c("dplyr",
             "purrr",
             "ggplot2", 
             "infer")

## Figure out which packages are not installed and install them
installed = vapply(packages, function(x) 
  length(find.package(x, quiet = TRUE)) > 0, 
  TRUE)

if(any(!installed)) {
  install.packages(packages[!installed])
}

## Load packages
done = lapply(packages, require, character.only = TRUE)

```

## Mystery machine #1
```{r}

machine_spec = function(zero, mu) {
  z = (zero - 1 + mu)/19
  y = mu - 20*z
  
  c(zero, y, z)
}

outputs = c(0, 1, 20)
probs = list(`mu = 0.2` = machine_spec(0.8, .3),
             `mu = 0.7, low variability` = machine_spec(0.4, .7), 
             `mu = 0.7, high variability` = machine_spec(0.9, .7),
             `mu = 0.9, high variability` = machine_spec(0.9, 0.9))

# lapply(probs, "%*%", outputs) ## get expected values

n = 1500
set.seed(40)
machine_df = map_dfr(probs, ~ data.frame(Result = sample(outputs, prob = ., size = n, replace = TRUE), Sample_Size = 1:n), .id = "Machine") %>% 
  group_by(Machine) %>% 
  mutate(Cumulative_Mean = cummean(Result))


single_attempt = ggplot(machine_df, aes(Sample_Size, Cumulative_Mean, 
                       group = Machine, color = Machine)) +  
  geom_line(size = 1.2) + theme(legend.position = "top") + scale_color_brewer(palette = "Set3")

single_attempt
```
## Calculate null rejection region

$$H_0: \mu = 0.90$$
$$H_A: \mu < 0.90$$
If anything, we expect the casinos to cheat the customers, not to favor them, so we only need a one-sided hypothesis.

We can determine where the rejection region would be using simulation, again, for n = 1500.

```{r}
n = 1500
K = 1000 # Number of trials
gen_trials = function(prob, n, K, outputs) {
  map_dbl(1:K, ~ mean(sample(outputs, prob = prob, size = n, replace = TRUE)))
}

alpha = 0.05
cutoff = alpha*1500

null_trials = data.frame(Mean_1500 = gen_trials(probs[[4]], n, K, outputs)) %>% 
  arrange(Mean_1500) %>%
  mutate(Below_Cutoff = (1:n()) < cutoff)
  
reject = max(null_trials$Mean_1500[null_trials$Below_Cutoff])

ggplot(null_trials, aes(Mean_1500, color = Below_Cutoff)) + 
  geom_dotplot(dotsize = 0.2) + scale_color_manual(values = c("red", "blue")) +
  geom_vline(xintercept = reject) + 
  theme(legend.position = "none") 
```


The bottom 5% of trials are below .748.  Now, we can calculate power for each of the scenarios above -- this is the proportion of trials where the mean is below 0.748.

```{r}
machine_trials = map_dfr(probs[-4], ~ data.frame(Mean_1500 = gen_trials(., n, K, outputs)), .id = "Machine") %>% 
  arrange(Mean_1500) %>%
  mutate(Rejected = Mean_1500 < reject)


ggplot(machine_trials, aes(Mean_1500, color = Rejected)) + 
  geom_dotplot(dotsize = 0.2) + scale_color_manual(values = c("red", "blue")) +
  geom_vline(xintercept = reject) + 
  facet_wrap(~ Machine) +
  theme(legend.position = "none") 

```
Power in each case
```{r}

with(machine_trials, table(Machine, Rejected)) %>% prop.table(margin = 1)

```

